---
title: "Setup Git Credentials"
author: "Quinn Zilly"
date: "2026-05-22"
output: html_document
---

1.  Delete hash tags and run to install the `usethis` & `gitcreds` package

```{r}
#install.packages("usethis")
#install.packages("gitcreds")
```

2.  Load the `usethis` & `gitcreds` packages

```{r, message=F, warning=F}
library(usethis)
library(gitcreds)
```

3.  Enter the username and email you used to sign up for GitHub, then run this code chunk. The `user_git_config()` function tells GitHub that this computer is associated with your account.

```{r}
usethis::use_git_config(
  user.name  = "Jane Doe",
  user.email = "jane@example.com"
)
```

4.  Run `create_github_token()` to generate a PAT.

```{r}
usethis::create_github_token()
```

This will direct you to the [GitHub Website](https://github.com/settings/tokens). You will have three options before generating the token:

1.  **Note** - This is to remind your future self what this key was for, as you may eventually have many keys for different machines or projects. For now, name it something like "*yourname-personal-pc*" to remind you that the key is for this computer.
2.  **Expiration** - GitHub recommends setting tokens to expire every 30 days for security purposes. If you set your token to expire, you will need to generate a new one in the future.
3.  **Select Scopes** - Scopes define what this token gives access to on GitHub. For now, accept the defaults (repo, workflow, gist, user).

**IMPORTANT**: Make sure to copy the key to your clipboard before you leave the page, as you will not be able to view it again! (you will have to delete the token and make a new one if you do)

5.  Run `gitcreds_set()`, then paste your token into the console when prompted.

```{r}
gitcreds::gitcreds_set()
```
