---
title: "7. Mediation and Bootstrapping"
output: html_document
date: "`r Sys.Date()`"
---

## Research Vignette
In this class for the class examples, I will be using data collected through the Emerging Adulthood project. This is a multi-site research experiment, with longitudinal waves being collected every few years. This is 2nd wave.

https://osf.io/te54b/

For specifics on the variables, the surveys can be found on the OSF site, and a quick data dictionary is available on Canvas with most of the key variables defined.

Today, I'm curious about the relationship between perceived social support from friends and stress, and if that relationship is mediated by a sense of belonging.

Predictor: PSS_Friends
Mediator: BelongingScale
Outcome: Stress


Some packages I frequently use:
```{r}
library(psych)
library(apaTables)
library(lavaan)
library(tidyverse)
library(pwr2ppl) #https://cran.r-project.org/web/packages/pwr2ppl/index.html
```


Read in the data and name the object.
```{r}
dat <- read.csv("EmergingAdulthoodData.csv")
```


## Mediation

Mediation is used to describe how effects occur. The mediator variable is often the mechanism through which the predictor variable affects the outcome.

```{r}
model <- mediate(Stress ~ PSS_Friends + (BelongingScale), data = dat)
model
```

## Bootstrapping
Bootstrapping is a process of sampling with replacement. It relies on the assumption that the sample is representative of the population.

https://www.r-bloggers.com/2022/12/how-to-perform-bootstrapping-in-r/ 

```{r}
animals <- c("cat", "dog", "koala", "hippo", "orca")
animals
sample(animals, length(animals), replace = TRUE)
```

## Use the lavaan package

This is the most flexible and my recommended approach for simplicity but it requires model syntax to specify the paths correctly.

```{r}
model <- ' Stress ~ c*PSS_Friends
             
             BelongingScale ~ a*PSS_Friends
             Stress ~ b*BelongingScale
           
           # indirect effect (a*b)
             ab := a*b
           
           # total effect
             total := c + (a*b)
         '

fit <- sem(model, se = "bootstrap", data = dat)
summary(fit)
```



## Process Macro

This is a flexible tool for mediation, moderation, and moderated mediation. It uses bootstrapping to calculate the indirect effect.

```{r}
# This can take a few minutes to run
source('process.R')
```
```{r}
process(data = dat, y = "Stress", x = "PSS_Friends", m = "BelongingScale", model = 4, boot = 5000, seed = 1234, progress=F)
```


## APA style RESULTS write-up

We used a simple mediation analysis to examine the relationship between perceived social support from friends and stress through a sense of belonging. Perceived social support from friends was found to be a statistically significant predictor of belongingness, $a = 0.05$, $p < .001$, 95% CI [0.03, 0.07]. Additionally, belongingness was found to be a statistically significant predictor of stress, $b = 0.30$, $p < .001$, 95% CI [0.27, 0.33]. Analyzing the indirect effect using 5000 bootstrapped samples, perceived social support from friends had a positive effect on stress indirectly through a sense of belongingness (estimate: 0.02, 95% CI [0.01, 0.02]).


## Power Analysis for Mediation
https://cran.r-project.org/web/packages/pwr2ppl/index.html

a1 path between predictor and first mediator 
a2 path between predictor and second mediator 
b1 Path between first mediator and dependent variable 
b2 Path between second mediator and dependent variable 
rm1m2 Correlation first mediator (m1) and second mediator (m2) 
cprime Path between predictor and dependent variable 
n Sample size 
alpha Type I error (default is .05) 
mvars Number of Mediators 
rep number of repetitions (1000 is default)


Run the power analysis
```{r}
medjs_paths(a1 = 0.0522, a2=NULL, b1 = 0.30, b2=NULL, rm1m2=NULL, n = 2850, cprime = -.12, alpha=0.05, mvars = 1, rep=1000)
```


### Note for HW

The data is in an Excel file, on one of multiple sheets. Students will need to use the library(readxl) package and dat <- read_excel("mydata.xlsx", sheet = "Sheet1")


