---
title: "6. Moderation (continuous)"
output: html_document
date: "`r Sys.Date()`"
---

## Research Vignette
In this class for the class examples, I will be using data collected through the Emerging Adulthood project. This is a multi-site research experiment, with longitudinal waves being collected every few years. This is 2nd wave.

https://osf.io/te54b/

For specifics on the variables, the surveys can be found on the OSF site, and a quick data dictionary is available on Canvas with most of the key variables defined.

Today, I'm curious about the relationship between mindfulness and stress. Additionally, I want to know that this relationship depends on a person's income. All the variables here are continuous.

Predictor: Mindfulness
Moderator: Income
Outcome: Stress


Some packages I frequently use:
```{r}
library(psych)
library(apaTables)
library(tidyverse)
library(lm.beta)
library(pwr2ppl)
library(interactions)
```


Read in the data and name the object.
```{r}
dat <- read.csv("../Data/EmergingAdulthoodData.csv")
```


## Moderation with a continuous moderator

Moderation is used when there is an interaction between predictor variables in the model. Statistically, nothing is included to differentiate which variable is the predictor and which is the moderator.

```{r}
model <- lm(Stress ~ MindfulnessScale * income, data=dat)
model
```

Assess statistical significance and report model fit via the proportion of variance explained.

```{r}
model %>% summary()
```

Confidence Intervals
```{r}
model %>% confint()
```

## Plot the Interaction
```{r}
interact_plot(model, pred = "MindfulnessScale", modx = "income", interval = TRUE)
```


## APA style RESULTS write-up

Moderation analysis was conducted to evaluate the extent to which income moderates the relationship between mindfulness and stress. There was a statistically significant interaction between mindfulness and income, $b = -0.02$, $p = .005$, $95% CI [-0.03, -0.01]$. The conditional effect of mindfulness on stress was found to be statistically significant as well, $b = -0.28$, $p < .001$, $95% CI [-0.33, -0.24]$, when income was at 0. However, the conditional effect of income on stress was not statistically significant, $b = -0.03$, $p = .10$, $95% CI [-0.01, 0.07]$, conditioned on mindfulness being 0. These predictors together explained a significant proportion of variance in stress scores, $R^2 = .20$, $F(3, 3078) = 252.90$, $p < .001$, indicating that the model where income moderates the relationship between mindfulness and stress explained approximately 20% of the variance in stress.


## Power Analysis for Moderated Multiple Linear Regression
https://cran.r-project.org/web/packages/pwr2ppl/index.html

R2Mod = Full Model R2
R2Ch = Change in R2 Added by Interaction
mod_pred = Full Model Number of Predictors
ch_pred = Change Model Number of Predictors
nlow = starting sample size
nhigh = ending sample size
by = incremental increase in sample(e.g. nlow=10, nhigh=24, by=2, produces estimates of 10, 12, and 14... 24)
alpha = Type I error (default is .05)


Get R2 value for the model without the interaction
```{r}
no_int_model <- lm(Stress ~ MindfulnessScale + income, data=dat) 
summary(no_int_model)$r.squared

# remind ourselves of R^2 from our interaction model
summary(model)$r.squared
```
Save the difference as an object
```{r}
changeR2 <- summary(model)$r.squared - summary(no_int_model)$r.squared
```


Run the power analysis
```{r}
regintR2(R2Mod=summary(model)$r.squared, R2Ch=changeR2, mod_pred=3, ch_pred=1, nlow=2000, nhigh=3000, by=100)
```





