---
title: "Descriptive Statistics In Class"
output: html_document
date: "`r Sys.Date()`"
---

Load Packages

```{r}
library(psych)
library(stats)
library(apaTables)
library(tidyverse)
```

```{r}
options(scipen = 999)
```


Read in the data and name it as an object

```{r}
dat <- read.csv("data/CleanCharacterDataStudy4a.csv")
```



## Central Tendency

First visualize the data. Let's make a histogram of commute. ChatGPT can help figure out what each line does and I love editing code from the R Graphics Gallery: https://r-graph-gallery.com/

```{r}
ggplot(dat, aes(x = CCC_students)) +
  geom_histogram(binwidth = 1, color = "black", fill = "skyblue") +
  theme_minimal() +
  labs(
    title = "Histogram of the CCC - Students Variable",
    x = "Number of Students",
    y = "Frequency"
  )
```


Central Tendency

```{r}
mean(dat$CCC_students)
```

Median of commute time (especially if there are outliers)
```{r}
median(dat$CCC_students)
```

Mode needs a categorical variable, like Gender. Also, no actual function in base R.
```{r}
table(dat$Gender)
```

## Variability

Variance
```{r}
var(dat$CCC_students)
```

Standard deviation, which is the square root of the variance.
```{r}
sd(dat$CCC_students)
sqrt(var(dat$CCC_students))
```

Check for normality (skew and kurtosis)
```{r}
psych::describe(dat$CCC_students)
```



Correlations
```{r}
cor.test(dat$CCC_students, dat$CCC_profs)
```

Making correlation tables in APA format
```{r}
continuousVars <- data.frame(dat$CCC_students, dat$CCC_profs, dat$Age)
apa.cor.table(continuousVars, filename = "cortable.doc")
```












