---
title: "One Sample Tests In Class"
output: html_document
date: "`r Sys.Date()`"
---

Load Packages
```{r}
library(tidyverse)
```

Chi-square test for the Taco Bell sauce

```{r}
# Example observed frequencies
observed <- c(8, 9, 7, 9)

# Expected proportions (must sum to 1)
expected_prop <- c(.25, .25, .25, .25)

# Run one-sample chi-square test
chisq.test(x = observed, p = expected_prop)
```

Load the Old Faithful geyser eruption dataset
```{r}
dat <- read.csv("OldFaithful.csv")
```


Look at the variables and choose what we want to test
```{r}
head(dat)
```

A tour company promises an eruption every hour. Is that true?

Null hypothesis: the time between eruptions is 60 minutes
Alternative hypothesis: the time between eruptions is NOT 60 minutes (two-tailed)

```{r}
mean(dat$waiting)
```

```{r}
psych::describe(dat$waiting)
```

Don't forget to change the null hypothesis
```{r}
t.test(dat$waiting, mu=60)
```
Our results are statistically significant. This one-sample t test showed that the sample average wait time between eruptions of 70.89 minutes at Old Faithful differ significantly from the hypothesized population mean of 60 minutes, $t(271) = 13.22$, $p < .001$. 


Try a one tailed test (greater, less, or two.sided by default). What changes? What stays the same?
```{r}
t.test(dat$waiting, mu=60, alternative = "less")
```

Visualize with a histogram
```{r}
ggplot(faithful, aes(x = waiting)) +
  geom_histogram(bins = 20, fill = "skyblue", color = "black") +
  geom_vline(xintercept = 60, linetype = "dashed") +
  labs(title = "Old Faithful Time Between Eruptions",
       x = "Wait Time (minutes)", y = "Count")
```


Now your turn with the GTAs or Dr. Fossum: do this same test for duration of the eruption.




















