---
title: "t Tests"
output: html_document
date: "`r Sys.Date()`"
---

Load packages and read in the data
```{r, message = F}
library(tidyverse)
library(psych)
```


Data for paired samples test
```{r}
dat <- read.csv("../Data/stroop.csv")
```

Calculate the change score
```{r}
dat$change <- dat$incongruent - dat$congruent
```

Get descriptive statistics
```{r}
describe(dat$congruent)
describe(dat$incongruent)
describe(dat$change)
```

Assumptions of a t test

Normality
```{r}
hist(dat$change)
```

Test the normality assumpution using the Shapiro Wilks test
```{r}
shapiro.test(dat$change)
```


Run a paired sample t test for the stroop data
```{r}
t.test(x = dat$incongruent, y = dat$congruent,
       mu = 0, paired = TRUE, var.equal = FALSE)
```
Same as one sample test for the change variable (it's the same!)

```{r}
t.test(dat$change)
```

Independent Sample t Test

Pretend we asked another question to separate students into two groups. This grouping variable is called GROUPS.

Run this code but ignore it. It's just making up data to do this independent samples t test to make the grouping variable.
```{r}
dat$groups <- c(0,0,0,0,0,0,0,0,1,1,1,1,1,1,1,1,1)
```

Run an independent samples t test using code that follows the same pattern as simple linear regression in the lm() command.
```{r}
t.test(control ~ groups, data = dat)
```

Magic of regression: run this as a linear model and see what is the same

-square the t statistic to get the F statistic
 
-coefficient is just the mean difference

-intercept is the mean of the reference group

```{r}
lm(control ~ groups, data = dat) %>% summary()
```










